---
title: "Home-Work-2"
author: "VigneshwarPesaru"
date: "3/19/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


## R Reading Libraries

```{r reading the libraries}
	message = FALSE
	warning = FALSE

library(tidyverse) # metapackage with lots of helpful functions
library(RColorBrewer)
library(scales)
library(lubridate)
library(car)
#options(warn=-1)
library(AppliedPredictiveModeling)
library(ggcorrplot)
library(lattice)
library(psych)
library(DataExplorer)
library(readxl)
library(data.table)
library(plotly)
library(e1071)
library(glmnet)
library(caret)
library(gridExtra)	
library(plyr)
#install.packages("ROCR")
library(ROCR)
	library(pROC)
	library(randomForest)
#install.packages('reticulate')
main_churn_data = read_csv('WA_Fn-UseC_-Telco-Customer-Churn.csv')
#The main churn data is the telecom data set



```


#Check the number of missing values in each column
```{r}
sapply(main_churn_data, function(x) sum(is.na(x)))
#removing the missing values if any
main_churn_data <- main_churn_data[complete.cases(main_churn_data), ]

```
```{r }

summary(main_churn_data)


```

Description:

###########################OUTCOME VARIABLE##############################

The potential outcome variable in the main churn data set is the
"churn." If the customer is already left from the service, then he/she is
given by churn "YES" else "NO"

# Summary Statistics
  
  Datatype : Binary (0-1)
  churn    : Yes(Customer already left the service because of several reasons)
  churn    : No(Customer is in the service program)
  
  
########The variables invloved in the main_churn_data process are:#######

1. CustomerID: This is the unique(primary key) is given to all the customers
who are currently in the service along with the customers who already left the 
service. This column is mainly used for east identification of any given customer.
  
     Datatype: character
     Length:7032

2. Gender: This column is just know the gender who are enrolled and left the telecom services
  
     Datatype: character
     Gender  : Male
     Gender  : Female
     
3. SeniorCitizen: This is the measurement taken to know whether the given citizen
is a senior citizen or not.
  
     Datatype: Binary (0-1)
     SeniorCitizen: 1 (if the given customer is senior citizen)
     SeniorCitizen: 0 (if the given customer is not a senior citizen)

4. Partner: Whether the customer has a partner or not (Yes, No)
  
     Datatype: Binary (0-1)
     Partner: 1 (if the given customer have a partner)
     Partner: 0 (if the given customer dont have a partner)

5. Dependent: Whether the customer has a dependents or not (Yes, No)
  
     Datatype: Binary (0-1)
     Partner: 1 (if the given customer have a dependent)
     Partner: 0 (if the given customer dont have a dependent)

6. Tenure:Number of months the customer has stayed with the company

     Datatype: Integer (Non-negative)
      -Min.   :1.00    
      -1st Qu.:9.00    
      -Median :29.00  
      -Mean   :32.42    
      -3rd Qu.:55.00    
      -Max.   :73.00  
      -Range  :[1, 72]  
   
        
7. PhoneService: Whether the customer has a phone service or not (Yes, No)
     
     Datatype: Binary (0-1)
     PhoneService: 1 (if the given customer have a PhoneService)
     PhoneService: 0 (if the given customer dont have a PhoneService)
   
        
8. MultipleLines : Whether the customer has multiple lines or not 
  (Yes, No, No phone service)
   
     Datatype: Binary (0-1)
     MultipleLines: 1 (if the given customer have a MultipleLines)
     MultipleLines: 0 (if the given customer dont have a MultipleLines)
   
        
9. InternetService  : Customer’s internet service provider 
   (DSL, Fiber optic, No)

     Datatype: character
     InternetService  : DSL(if the customer have DSL)
     InternetService  : Fiber optic(if the customer have a Fiber optic)
     InternetService  : No(if the customer dont have any)
     
        
10. OnlineSecurity  : Whether the customer has online security or not (Yes, No, No internet service)
     
     Datatype: Binary (0-1)
     OnlineSecurity  : Yes (if the customer have online security)
     OnlineSecurity  : No (if the customer dont have an online security).

11. OnlineBackup  : Whether the customer has online backup or not (Yes, No, No internet service)
     
     Datatype: Binary (0-1)
     OnlineBackup  : Yes (if the customer have OnlineBackup )
     OnlineBackup  : No (if the customer dont have an OnlineBackup).

12. DeviceProtection  : Whether the customer has device protection or not (Yes, No, No internet service)

     Datatype: Binary (0-1)
     DeviceProtection  : Yes (if the customer have DeviceProtection )
     DeviceProtection  : No (if the customer dont have an DeviceProtection).
    
13. Techsupport  :Whether the customer has tech support or not (Yes, No, No internet service)

     Datatype: Binary (0-1)
     Techsupport  : Yes (if the customer have Techsupport )
     Techsupport  : No (if the customer dont have an Techsupport).
     
14. Streaming Tv  : Whether the customer has streaming TV or not (Yes, No, No internet service)

     Datatype: Binary (0-1)
     streamingtv  : Yes (if the customer have streamingtv )
     streamingtv  : No (if the customer dont have an streamingtv).     
    
    
15. Streaming Movies  : Whether the customer has streaming movie or not (Yes, No, No internet service)

     Datatype: Binary (0-1)
     streamingMovie  : Yes (if the customer have streamingmovie )
     streamingMovie  : No (if the customer dont have an streaming movie).      
    
16. Contract  : The contract term of the customer (Month-to-month, One year, Two year)

     Datatype: character
     Contract  : Month-to-month
     Contract  : One Year
     Contract  : Two Year
     
16. PaperlessBilling  : Whether the customer has paperless billing or not (Yes, No)

     Datatype: Binary (0-1)
     PaperlessBilling :  Yes (if the customer have paperless billing )
     PaperlessBilling :  No (if the customer dont have paperless billing )


17. MonthlyCharges:Number of months the customer has stayed with the company

     Datatype: Real (Non-negative)
      -Min.   :18.25    
      -1st Qu.:35.59   
      -Median :70.35  
      -Mean   :64.80    
      -3rd Qu.:89.86    
      -Max.   :118.75 
      -Range  :[18.25, 118.75]
      
        
17. TotalCharges:The total amount charged to the customer

     Datatype: Real (Non-negative)
      -Min.   :18.8    
      -1st Qu.:401.4   
      -Median :1397.5  
      -Mean   :2283.5    
      -3rd Qu.:3794.7    
      -Max.   :8684.4
      -Range  :[18.8, 8684.4]

18. Churn: Whether the customer churned or not (Yes or No)
     
     Datatype: Binary (0-1)
     Churn :  Yes (if the customer churned )
     Churn :  No (if the customer not churned )
    
```{r DataPreprocessing }
message = FALSE
warning = FALSE

main_churn_data$SeniorCitizen <- as.factor(mapvalues(main_churn_data$SeniorCitizen,
                                          from=c("0","1"),
                                          to=c("No", "Yes")))

#main_churn_data$Churn <- #as.factor(mapvalues(main_churn_data$Churn,
#                                          from=c("No","Yes"),
#                                          to=c("0", "1")))


main_churn_data$MultipleLines <- as.factor(mapvalues(main_churn_data$MultipleLines, 
                                           from=c("No phone service"),
                                           to=c("No")))


main_churn_data$OnlineBackup <- as.factor(mapvalues(main_churn_data$OnlineBackup, 
                                           from=c("No internet service"),
                                           to=c("No")))


main_churn_data$OnlineSecurity = as.factor(mapvalues(main_churn_data$OnlineSecurity,
                                           from=c("No internet service"),
                                           to=c("No")))



main_churn_data$DeviceProtection = as.factor(mapvalues(main_churn_data$DeviceProtection,
                                           from=c("No internet service"),
                                           to=c("No")))


main_churn_data$TechSupport = as.factor(mapvalues(main_churn_data$TechSupport,
                                           from=c("No internet service"),
                                           to=c("No")))

main_churn_data$StreamingTV = as.factor(mapvalues(main_churn_data$StreamingTV,
                                           from=c("No internet service"),
                                           to=c("No")))

main_churn_data$StreamingMovies = as.factor(mapvalues(main_churn_data$StreamingMovies,
                                           from=c("No internet service"),
                                           to=c("No")))



main_churn_data$customerID <- NULL

temp_churn = main_churn_data



library(dplyr)

temp_churn <- temp_churn %>%
mutate_if(is.factor, as.numeric)

temp_churn$gender <- ifelse(temp_churn$gender == "Male",1,0)
temp_churn$Partner <- ifelse(temp_churn$Partner == "Yes",1,0)
temp_churn$Dependents <- ifelse(temp_churn$Dependents == "Yes",1,0)
temp_churn$PhoneService <- ifelse(temp_churn$PhoneService == "Yes",1,0)
temp_churn$PaperlessBilling <- ifelse(temp_churn$PaperlessBilling == "Yes",1,0)

type_of_service <- c("DSL" = 1, "Fiber optic" = "2", "No" = 2)

Internetservice <- factor(unname(type_of_service[temp_churn$InternetService]))

temp_churn$InternetService = NULL

temp_churn$Contract =  NULL

temp_churn$PaymentMethod =  NULL

temp_churn = cbind(Internetservice, temp_churn)

```   
   
```{r skewness and density of the outcome variable  }
#Gender plot
p1 <- ggplot(main_churn_data, aes(x = gender)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Senior citizen plot
p2 <- ggplot(main_churn_data, aes(x = SeniorCitizen)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Partner plot
p3 <- ggplot(main_churn_data, aes(x = Partner)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Dependents plot
p4 <- ggplot(main_churn_data, aes(x = Dependents)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Plot demographic data within a grid

grid.arrange(p1, p2, p3, p4, ncol=2)

```


   
   

```{r }
message = FALSE
warning = FALSE
#Phone service plot
p5 <- ggplot(main_churn_data, aes(x = PhoneService)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Multiple phone lines plot
p6 <- ggplot(main_churn_data, aes(x = MultipleLines)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Internet service plot
p7 <- ggplot(main_churn_data, aes(x = InternetService)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Online security service plot
p8 <- ggplot(main_churn_data, aes(x = OnlineSecurity)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Online backup service plot
p9 <- ggplot(main_churn_data, aes(x = OnlineBackup)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Device Protection service plot
p10 <- ggplot(main_churn_data, aes(x = DeviceProtection)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Tech Support service plot
p11 <- ggplot(main_churn_data, aes(x = TechSupport)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Streaming TV service plot
p12 <- ggplot(main_churn_data, aes(x = StreamingTV)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Streaming Movies service plot
p13 <- ggplot(main_churn_data, aes(x = StreamingMovies)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

#Plot service data within a grid
grid.arrange(p5, p6, p7,
             p8, p9, p10,
             p11, p12, p13,
             ncol=3)

```


```{r }

par(mfrow=c(2,2))
## if you save your hist() , it actually stores each bin count,midpoint of each stack,the density of each stack , etc..
## can use text fucntion to add text to plots
b<-hist(main_churn_data$tenure,xlim=c(0,80),breaks=10,main='Tenure freq histo',ylab='freq',xlab='Tenure',col=2)
text(b$mids,b$counts,labels=b$counts, adj=c(0.5, 1))

e<-hist(main_churn_data$TotalCharges,xlim=c(0,10000),breaks=8,main='Total Charges freq histo',ylab='freq',xlab='Total Charges',col=3 )
text(e$mids,e$counts,labels=e$counts, adj=c(0.5,1))

t<-hist(main_churn_data$MonthlyCharges,xlim=c(0,120),breaks=8, main='Monthly charges freq histo',ylab='freq',xlab='Monthly charges',col=5)
text(t$mids,t$counts,labels=t$counts, adj=c(0.5, 1))


```
```{r }
p20 <- ggplot(main_churn_data, aes(x = Churn)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)
p20
```

```{r }
message = FALSE
warning = FALSE
set.seed(123)


training.samples <- temp_churn$Churn %>%
  createDataPartition(p = 0.8, list = FALSE)

train.data  <- temp_churn[training.samples, ]
test.data <- temp_churn[-training.samples, ]




lda_model <- train(Churn ~., data = train.data,
                 method = "glm", 
                 trControl = trainControl("cv", number = 10, classProbs = TRUE,summaryFunction=twoClassSummary), 
                 metric = "ROC")

lda_pred <- predict(lda_model, test.data)
confusionMatrix(lda_pred, as.factor(test.data$Churn), positive = "No")


probs <- predict(lda_model, test.data,
type = "prob")[, "No"]



test.data$Churn = ifelse(test.data$Churn == "Yes", 1, 0)

#plot(roc(test.data$Churn, lda_pred$posterior[,"1"], direction="<"),
#     col="yellow", lwd=3, main="The turtle finds its way")

roc_lda=roc(response=test.data$Churn, predictor= factor(probs, 
ordered = TRUE), plot=TRUE)
plot(roc_lda, col="red", lwd=3, main="ROC curve Logistic",asp = NA)
#auc_lda<-auc(roc_lda,"auc")

```



```{r Penalized model }


set.seed(100)
ctrl2 <- trainControl(method = "cv", number = 10,
# extra lines required for ROC
classProbs = TRUE,
summaryFunction=twoClassSummary)
set.seed(100)
glmnGrid <- expand.grid(lambda = seq(0.0, 0.02, length = 10),
alpha = seq(0.0, 1, length = 10))
glmnTuned2 <- train(Churn ~ .,
                    data = train.data,
                    method = "glmnet",
                    tuneGrid = glmnGrid,
                    preProc = c("center", "scale"),
                    metric="ROC", # Switch metric
                    trControl = ctrl2)

glmnTuned2$results %>% filter(alpha == glmnTuned2$bestTune$alpha,
lambda == glmnTuned2$bestTune$lambda)

test.data$Churn = ifelse(test.data$Churn == "1", "Yes", "No")
glmnTuned2_pred <- predict(glmnTuned2, test.data)
confusionMatrix(glmnTuned2_pred, as.factor(test.data$Churn), positive = "No")


probs <- predict(glmnTuned2, test.data,
type = "prob")[, "No"]



roc_glmTuned2=roc(response=test.data$Churn, predictor= factor(probs, 
ordered = TRUE), plot=TRUE)
plot(roc_glmTuned2, col="red", lwd=3, main="ROC curve Penalized",asp = NA)
```





```{r Random Forest}
# Fit the model on the training set
ctrl = trainControl(method = "cv", number=10, 
                classProbs = TRUE, summaryFunction = twoClassSummary)

model.rf = train(Churn ~., data = train.data ,
                         method = "rf",
                         ntree = 75,
                         tuneLength = 5,
                         metric = "ROC",
                         trControl = ctrl)
        

model.rf

rf_pred <- predict(model.rf, test.data)

#confusionMatrix(rf_pred, as.factor(test.data$Churn), positive = "No")

rf_probs <- predict(model.rf, test.data,
type = "prob")[, "No"]

confusionMatrix(rf_pred, as.factor(test.data$Churn), positive = "No")

roc_rf=roc(response=test.data$Churn, predictor= factor(rf_probs, 
ordered = TRUE), plot=TRUE)
plot(roc_rf, col="red", lwd=3, main="ROC curve RF",asp = NA)
#auc_lda<-auc(roc_rf)


```

```{r naive bayes model}
# Fit the model on the training set
ctrl = trainControl(method = "cv", number=10, 
                classProbs = TRUE, summaryFunction = twoClassSummary)

model.nb = train(Churn ~., data = train.data ,
                         method = "nb",
                         trControl = ctrl)
        

model.nb

nb_pred <- predict(model.nb, test.data)
#confusionMatrix(nb_pred, as.factor(test.data$Churn), positive = "No")

nb_probs <- predict(model.nb, test.data,
type = "prob")[, "No"]



roc_nb=roc(response=test.data$Churn, predictor= factor(nb_probs, 
ordered = TRUE), plot=TRUE)
plot(roc_nb, col="red", lwd=3, main="ROC curve Naive",asp = NA)
#auc_nb<-auc(roc_nb)


```

# Comparing the models with repect to RMSE and Model compexities.



1. Linear Model:

    In the case of Linear regression model the estimated RSME is 1.21.

2. LASSO Model:

    2.1.   In the case of LASSO model the estimated RSME is 1.18, 
    
    2.2.  From the graph of "Fraction of Full solution" we can observe the following:
   
    2.3:  RMSE value started decreasing as the Fraction of Full solution increases
          
   
         Fraction Full Solution is give by:
         
           * (LASSO Solution)/(Linear regression Solution)
           
        Note: Ultimately between both the models(Linear and LASSO) the LASSO performed
        well but during the starting phase of iterations from the graph the LASSO 
        performed poor compared to the Linear regression model.

3. Regression Tree:

    3.1. The estimated RSME value in the case of regression Tree is 1.86. 
    
    3.2. Clearly from the regression tree estimated RMSE value, we can conclude that regression
         trees performed very poor compared to rest of the models.
        
    3.2. This could probably the regression trees are overfitting the data by
         splitting the predictors space in illogical way. Hence we end up with
         higher estimated RSME.
         
    3.3. From the graph, we can clearly seen that as the complexity parameter(cp)
         increases the RSME values started increasing. 
         
         Note: Although the model has considered the optimal cp value for which 
         the RSME value is low but on the test set it has performed very poor.
         
         Hence we can drop the idea of implementing the Regression Tree model 
         for this Data set. 
         
4. Neural Network:

    4.1 Simple Neural Network:
    
        * The estimated  Training RMSE value for the simple Neural Network is
          1.10 which is lower than the test estimated RMSE which is 2.10.
          
          Clearly the simple neural networks has overfitted our model.  
        
    Lets see how we have fixed the overfitted simple neural network.    
        
    4.2 Averaging Neural Network:
    
        * The estimated RMSE value for the Averaging Neural Network is 0.96 which 
        is lower than the rest of the models.
        
        * The primary reason why the averaging neural network worked out well 
          when compared to simple neural network is because initialization step 
          in Neural Network plays a impactful role in prediction.
          
        * In the case of averaging, we choose three random initializations, then
          we try to fit the model based on each initialization by calling 
          f1(x), f2(x), f3(x).
          
        * Based on this we try to predict for 1/3*(f1(x)+f2(x)+f3(x)), in this
          way we can escape from overfitting.
          
     
     Therefore, till here we can say that:
      
      RMSE(AverageNeuralNet) << RMSE(LASSO) << RMSE(Linear Regression) << RMSE(Regression Tree)
      
      
      
5. Support Vector Mission(SVM):

      5.1. The estimate RSME value with the SVM is low compared to Neural Network.
      
      5.2. The cost or loss function is chosen to be 8 for this data, which is optimal.
           The loss function really helped in improving the model by reducing the overall RMSE to a very lower number.
           
      5.3. Although this is a slight overfit model, but this model has performed
           well when compared to Average Neural Network.
           
           
Conclusion:
           
     * Hence out of all the models, based on the RMSE and complexity, we can 
     now conclude that SVM out performed when compared to rest of the models.
     
     
      RMSE(SVM) << RMSE(AverageNeuralNet) << RMSE(LASSO) << RMSE(Linear Regression) << RMSE(Regression Tree)
      
      
      


        
         















